### 2022-01-15 manuscript/manuscript.Rmd
Update manuscript.Rmd

author:
  - name: 'Tom E. Hardwicke'
    affiliation: '1'
    corresponding: yes
    email: 'tom.hardwicke@uva.nl'
    address: 'Department of Psychology, University of Amsterdam, Nieuwe Achtergracht 129-B, 1018 WT Amsterdam'
  - name: 'Robert T. Thibault'
    affiliation: '2,8'
  - name: 'Jessica E. Kosie'
    affiliation: '4'
  - name: 'Loukia Tzavella'
    affiliation: '7'
  - name: 'Theiss Bendixen'
    affiliation: '3'
  - name: 'Vivian E. Köneke'
    affiliation: '5'
  - name: 'John P. A. Ioannidis'
    affiliation: '6,9'

affiliation:
  - id: '1'
    institution: 'Department of Psychology, University of Amsterdam'
  - id: '2'
    institution: 'Meta-Research Innovation Center at Stanford (METRICS), Stanford University'
  - id: '3'
    institution: 'Department of the Study of Religion, Aarhus University'
  - id: '4'
    institution: 'Department of Psychology, Princeton University'
  - id: '5'
    institution: 'Charité – Universitätsmedizin Berlin'
  - id: '6'
    institution: 'Meta-Research Innovation Center at Stanford (METRICS)'
  - id: '7'
    institution: 'School of Psychology, Cardiff University'
  - id: '8'
    institution: 'School of Psychological Science, Univeristy of Bristol'
  - id: '9'
    institution: 'Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Berlin Institute of Health, Charité – Universitätsmedizin Berlin AND Departments of Medicine, Epidemiology and Population Health, Biomedical Data Science, and Statistics, Stanford University'

@@

There was substantial variation across scientific domains with journals in Clinical Medicine standing out as offering the most PPPR options and publishing the most PPPR, but also imposed the most restrictive length and time-to-submit limits (for concordant evidence, see Altman, 2002; Von Elm et al., 2009; Winker, 2013). Other health-related domains seemed to have a more active PPPR culture than non-health related domains like the physical sciences and social sciences. Many domains, such as mathematics, showed little evidence of any PPPR activity, with few journals offering PPPR and scarce evidence of published PPPR. Our data do not speak to the causal forces that underlie these inter-domain differences, but some potential contributing factors could be cultural (e.g., different attitudes towards scientific criticism and how it should be handled), pragmatic (e.g., differences in methodological standards and research quality, manifesting in differential need for scientific criticism), bureaucratic (e.g., different resources available to support PPPR), or historic (e.g., individuals or events that highlighted the value of PPPR).

@@

Our exploration of how PPPR is used in practice suggested that letters are far more common than other types of PPPR, perhaps because they are the most frequently available PPPR option and also because, as formal articles, they may impart greater academic credit to their authors than informal web comments (Table 3). We found that half of the PPPRs we examined were behind a paywall, sometimes even when the target article was publicly accessible. This reduces access to PPPR for both professional scientists and other readers, like patients, journalists, and policy-makers (Winker, 2015). Most PPPR had conflict of interest statements and a third of those statements declared a potential conflict. Conflict of interest statements enable readers to evaluate an important risk of bias and seem just as relevant to PPPR as they are to other academic articles (Dunn et al., 2016).

The PPPRs we examined addressed a range of issues spanning the timeline of a research project, including design, implementation, analysis, reporting, and interpretation. The vast majority of PPPRs did not include new analysis of original or new data. This may be because very few original articles stated that data was available, as is typical in many scientific domains (Hardwicke et al., 2021; Hardwicke, Wallach, et al., 2020; Serghiou et al., 2021). Most of the PPPR was short (~250 words) and published within five months of the target article, perhaps partly because they were published in some journals that imposed the strictest limits on PPPR (Table 2).

>>>>

author:
  - name: 'Tom E. Hardwicke'
    affiliation: '1'
    corresponding: yes
    email: 'tom.hardwicke@uva.nl'
    address: 'Department of Psychology, University of Amsterdam, Nieuwe Achtergracht 129-B, 1018 WT Amsterdam'
  - name: 'Robert T. Thibault'
    affiliation: '2,8'
  - name: 'Jessica E. Kosie'
    affiliation: '4'
  - name: 'Loukia Tzavella'
    affiliation: '7'
  - name: 'Theiss Bendixen'
    affiliation: '3'
  - name: 'Sarah A. Handcock'
    affiliation: '10'
  - name: 'Vivian E. Köneke'
    affiliation: '5'
  - name: 'John P. A. Ioannidis'
    affiliation: '6,9'

affiliation:
  - id: '1'
    institution: 'Department of Psychology, University of Amsterdam'
  - id: '2'
    institution: 'Meta-Research Innovation Center at Stanford (METRICS), Stanford University'
  - id: '3'
    institution: 'Department of the Study of Religion, Aarhus University'
  - id: '4'
    institution: 'Department of Psychology, Princeton University'
  - id: '5'
    institution: 'Charité – Universitätsmedizin Berlin'
  - id: '6'
    institution: 'Meta-Research Innovation Center at Stanford (METRICS)'
  - id: '7'
    institution: 'School of Psychology, Cardiff University'
  - id: '8'
    institution: 'School of Psychological Science, Univeristy of Bristol'
  - id: '9'
    institution: 'Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Berlin Institute of Health, Charité – Universitätsmedizin Berlin AND Departments of Medicine, Epidemiology and Population Health, Biomedical Data Science, and Statistics, Stanford University'
  - id: '10'
    institution: 'Univeristy of Melbourne'

@@

There was substantial variation across scientific domains with journals in Clinical Medicine standing out as offering the most PPPR options and publishing the most PPPR, but also imposed the most restrictive length and time-to-submit limits (for concordant evidence, see Altman, 2002; Von Elm et al., 2009; Winker, 2013). In general, health-related domains seemed to have a more active PPPR culture than non-health related domains like the physical sciences and social sciences. Many domains, such as mathematics, showed little evidence of any PPPR activity, with few journals offering PPPR and scarce evidence of published PPPR. Our data do not speak to the causal forces that underlie these inter-domain differences, but some potential contributing factors could be cultural (e.g., different attitudes towards scientific criticism and how it should be handled), pragmatic (e.g., differences in methodological standards and research quality, manifesting in differential need for scientific criticism), bureaucratic (e.g., different resources available to support PPPR), or historic (e.g., individuals or events that highlighted the value of PPPR).

@@

Our exploration of how PPPR is used in practice suggested that letters are far more common than other types of PPPR, perhaps because they are the most frequently available PPPR option and also because, as formal articles, they may impart greater academic credit to their authors than informal web comments (Table 3). We found that half of the PPPRs we examined were behind a paywall, sometimes even when the target article was publicly accessible. This reduces access to PPPR for both professional scientists and other readers, like patients, journalists, and policy-makers (Winker, 2015). Most PPPR had conflict of interest statements and a third of those statements declared a potential conflict. Conflict of interest statements enable readers to evaluate an important risk of bias and seem just as relevant to PPPR as they are to other academic articles (Dunn et al., 2016). 
The PPPRs we examined addressed a range of issues spanning the timeline of a research project, including design, implementation, analysis, reporting, and interpretation. The vast majority of PPPRs did not include new analysis of original or new data. This may be because very few original articles stated that data were available, as is typical in many scientific domains (Hardwicke et al., 2021; Hardwicke, Wallach, et al., 2020; Serghiou et al., 2021). Most of the PPPR was short (~250 words) and published within five months of the target article, perhaps partly because they were published in some journals that imposed the strictest limits on PPPR (Table 2).
